### 2014-06-04 findsuperbase/findsuperbase.tex
added lib and data

\documentclass[a4paper,10pt]{article}

\usepackage{../mathbf-abbrevs}

\input{defs}

\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{calc}

%opening
\title{The number of lattices of Voronoi's first kind}
\author{Robby McKilliam}
\date{\today}

\begin{document}

\maketitle

% This document describes an algorithm to determine whether a lattice, specified by a basis matrix, is of Voronoi's first kind.  The algorithm returns an obtuse superbase for the lattice if the lattice is of Vornoi's first kind.

% An $n$-dimensional lattice $\Lambda$ is said to be of \emph{Voronoi's first kind} if it has what is called an \emph{obtuse superbasis}~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  That is, there exists a set of $n+1$ vectors $b_1,\dots,b_{n+1}$ such that $b_1,\dots,b_n$ are a basis for $\Lambda$,
% \begin{equation}\label{eq:superbasecond}
% b_1 + b_2 \dots + b_{n+1} = 0
% \end{equation}
% (the \emph{superbasis} condition), and the inner products satisfy
% \begin{equation}\label{eq:obtusecond}
% q_{ij} = b_i \cdot b_j \leq 0, \qquad \text{for} \qquad i,j = 1,\dots,n+1, i \neq j
% \end{equation}
% (the \emph{obtuse} condition).  The $q_{ij}$ are called the \emph{Selling parameters}~\cite{Selling1874}.  It is known that all lattices in dimensions less than $4$ are of Voronoi's first kind~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  An interesting property of lattices of Voronoi's first kind is that their relevant vectors have a straightforward description.

% \begin{theorem} \label{thm:revvecssuperbase} (Conway and Sloane~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases})
% The relevant vectors of $\Lambda$ are of the form,
% \[
% \sum_{i \in I} b_i
% \]
% where $I$ is a strict subset of $\{1, 2, \dots, n+1\}$ that is not empty, i.e., $I \subset \{1, 2, \dots, n+1\}$ and $I \neq \emptyset$.
% \end{theorem}

% Notes:
% \begin{itemize}
% \item Need definition of `strict' and `lax' relevant vectors. 
% \item How many lax relevant vectors can there be (presumably this comes the integer lattice, so $3^{n}-1$?)
% \item Are vectors from a superbase always strict?  Can probably show this using the arguments in Conway and Sloane's paper.
% \end{itemize}

% Our algorithm begins by enumerating a superset $R$ of the set of relevant vectors $\relevant(\Lambda)$.  This can be achieved using the algorithm of Micciancio and Voulgaris~\cite{Micciancio09adeterministic} or the algorithm described by Agrell~et.~al.~\cite[Sec.~VI.C]{Agrell2002}.  In many cases these algorithms will return precisely the set of relevant vectors and so $R = \relevant(\Lambda)$, but for what follows it is sufficient if $R \supseteq \relevant(\Lambda)$.

% If $\Lambda$ is of first kind then, by Theorem~\ref{thm:revvecssuperbase}, the vectors $b_1,\dots,b_{n+1}$ from the obtuse superbase are contained in $R$.  A straightforward procedure to determine whether the lattice is of first kind consists of enumerating all $\binom{\abs{R}}{n+1}$ combinations of $n+1$ vectors from $R$ and checking whether any combination forms a superbase.  That is, let $v_1,\dots,v_{n+1}$ be a set of $n+1$ distinct vectors from $R$.  If both $v_1+\dots + v_{n+1} = 0$ and $\dotprod{v_{i}}{v_{j}} \leq 0$ for all $i \neq j$ then the lattice is of first kind with superbase is $v_1,\dots,v_{n+1}$.  If no such combination of vectors exists in $R$, then the lattice is not of first kind.

% In general there are as many as $2^{n+1}-2$ relevant vectors (BLERG strict, lax etc?)

The plot shows the estimated (after 500 trials) probability of obtaining a lattice of Voronoi's first kind for $n=1,2,3,4$.  Three different methods for generating the $n \times n$ basis matrix $B$ are considered: 
\begin{enumerate}
\item entries drawn independently and uniformly in the interval $[0,1]$, 
\item entries drawn independently and Gaussian with zero mean and variance one, and 
\item the $i,j$th entry satisfies
\[
B_{i,j} = \begin{cases}
R_{i,j} & 1 \leq i,j \leq h \\
-C_{i-h,j} & h < i \leq n, 1 \leq j \leq h \\
C_{i,j-h} & h < j \leq n, 1 \leq i \leq h \\
R_{i-h,j-h} & h < i,j \leq n,
\end{cases}
\]
where $h = \ceil{n/2}$ is the smallest integer greater than or equal to $n/2$ and $R_{i,j}$ and $C_{i,j}$ for $i,j = 1,\dots,h$ are independently Gaussian distributed with zero mean and variance one. 
\end{enumerate}
When $n$ is even, the third case corresponds with the model commonly used for MIMO channels.  In this case $R$ and $C$ are $\tfrac{n}{2} \times \tfrac{n}{2}$ matrices containing the real and imaginary parts of the channel coefficients.

As expected, all lattices generated in dimensions 1, 2, and 3 were found to be of Voronoi's first kind.  When $n=4$ roughtly 25\% of the lattices generated were of Voronoi's first kind. This was the case for all 3 basis generation methods.  These simulations suggest that the lattices of Voronoi's first kind do not lie within a space of measure zero (in some sense!) in dimension 4.

I was too oppimistic when I thought I would be able to run this with $n=6$.  This would take too long with the existing code.  There are a number of optimisations that can be made that might make running simulations with $n=6$ possible.

\begin{figure}[tp]
\centering 
  \begin{tikzpicture}
    \begin{axis}[font=\footnotesize,clip mode=individual,height=7cm,width=9cm,legend style={draw=none,fill=none,legend pos=south west,font=\footnotesize}]    
      \addplot[mark=*,color=red,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/uniform};
      \addplot[mark=*,color=blue,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/gaussian};
      \addplot[mark=x,color=green,only marks,/pgfplots/error bars/.cd,
        x dir=none,
        y dir=both,
        y explicit] table[x=n,y=proportion,y error=stddev] {code/data/mimo};
\legend{Uniform, Gaussian}
    \end{axis}
  \end{tikzpicture}
\caption{Estimated probability of generating a lattices of Voronoi's first kind for different basis distributions and dimension $n=1,2,3,4$.}
\end{figure}

%\bibliography{../bib}

\end{document}