### 2014-08-26 reviews/proofedits.tex
added proofedits

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{../paper}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Finding a closest point in a lattice of Voronoi's first kind}
\author{Robby~G.~McKilliam, Alex Grant and I. Vaughan L. Clarkson
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

\begin{enumerate}

\item Page 1, AMS subject classifications: 11Y16

\item Page 2, paragraph 4, ``In this problem some ``advice'' about the lattice polynomial size is assumed to be given.''  This should be modified to ``In this problem some ``advice'' about the lattice is assumed to be given.''  That is, remove ``polynomial size''.

\item Page 3, the equation
\[
\rho = \frac{\|s\|}{2} = \frac{1}{2} \min_{x \in \Lambda / \{\zerobf\} } \|x\|
\]
The forward slash in $x \in \Lambda / \{\zerobf\}$ should be backslash like $x \in \Lambda \backslash \{\zerobf\}$.  The equation should appear as
\[
\rho = \frac{\|s\|}{2} = \frac{1}{2} \min_{x \in \Lambda \backslash \{\zerobf\} } \|x\|
\]

\item Page 14.  It is written:

\vspace{0.5cm}
\hspace{1cm}\parbox{10cm}{
The weight can now be written as
\begin{align*}
W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_i (1 - t_j) = F(t),
\end{align*}
for example. Finding a minimum \dots
}
\vspace{0.5cm}

Please remove ``for example'' and end the equation with a full stop.  The text should appear as

\vspace{0.5cm}
\hspace{1cm}\parbox{10cm}{
The weight can now be written as
\begin{align*}
W(C,\bar{C}) = \sum_{i \in C} \sum_{j \in \bar{C}} w_{ij} = \sum_{i =0}^{n+2} \sum_{j =0}^{n+2} w_{ij} t_i (1 - t_j) = F(t).
\end{align*}
Finding a minimum \dots
}
\vspace{0.5cm}

\end{enumerate}

\end{document}